---
title: "Cleaning Activity Data"
author: "S. Hughes"
date: "1/9/2017"
output: 
    md_document:
        variant: markdown_github
---

Created variables:
url                     - url of data sourced  
activity_labels         - vector of different activity names  
features                - vector of differnt feature names (X data)  
X_train/X_test          - Imported raw data for training/test sets  
subject_train/  
subject_tests           - Imported subjects id for training/test sets  
y_train/y_test          - Imported raw activiies for training/test sets  
y_train_labelled/  
y_test_labelled         - y values given descriptive labels  
training/test           - full data set for trainin/test (y, subj, X)  
completeData            - merged training/test data set  
X_meansd                - X data set with only "mean" or "std" features  
meansdData              - mean of X_meansd by activity and subject  
  
Downloads libraries used for markdown
knitr, tidyr, dplyr
```{r setup, max.print = 0}
library(knitr)
library(tidyr)
library(dplyr)
```

Downloads file from source
```{r getData}
temp <- tempfile()
url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
download.file(url, temp, method = "curl")
```

Downloads information about features and activity
```{r getLabels}
con <- unz(temp,"UCI\ HAR\ Dataset/activity_labels.txt")
activity_labels <- read.table(con)
activity_labels <- as.character(activity_labels$V2)

con <- unz(temp,"UCI\ HAR\ Dataset/features.txt")
features <- read.table(con)
features <- as.character(features$V2) 
```


Unzips and reads in training data
```{r getTraining}
con <- unz(temp,"UCI\ HAR\ Dataset/train/X_train.txt")
X_train <- read.table(con)

con <- unz(temp,"UCI\ HAR\ Dataset/train/y_train.txt")
y_train <- read.table(con)

con <- unz(temp,"UCI\ HAR\ Dataset/train/subject_train.txt")
subject_train <- read.table(con)
```

Unzips and reads in test data
```{r getTest}
con <- unz(temp,"UCI\ HAR\ Dataset/test/X_test.txt")
X_test <- read.table(con)

con <- unz(temp,"UCI\ HAR\ Dataset/test/y_test.txt")
y_test <- read.table(con)

con <- unz(temp,"UCI\ HAR\ Dataset/test/subject_test.txt")
subject_test <- read.table(con)

unlink(temp)
```

###3. Uses descriptive activity names to name the activities in the data set
Renames the y values by the description given in the activity labels
```{r label_y}
label_activity <- function(x) activity_labels[x]

y_train_labelled <- sapply(y_train, label_activity)
y_test_labelled <- sapply(y_test, label_activity)

table(y_train_labelled)
```

###4. Appropriately labels the data set with descriptive variable names.
Makes the training data into a single data frame. 
Renames the features by the description given in the features labels.
```{r makeTrain}
training <- data.frame(y_train_labelled, subject_train, X_train)
colnames(training) <- c("Activity", "Subject", features)
```
Makes the test data into a single data frame. 
Renames the features by the description given in the features labels.
```{r makeTest}
test <- data.frame(y_test_labelled, subject_test, X_test)
colnames(test) <- c("Activity", "Subject", features)
```


###1. Merges the training and the test sets to create one data set.
Binds test data and training data to get complete data
```{r makeComplete}
completeData <- rbind(training, test)
head(completeData[,1:4])
```

###2. Extracts only the measurements on the mean and standard deviation for each measurement.
Extracts all the features with either mean or standard deviation included in the discription. This includes mean frequency.
```{r meansdMeasures}
X_meansd <- completeData[,grepl("mean|std",names(completeData))]
head(X_meansd[,1:4])
```

###6. From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject

Prints out the mean of each variable for each activity (6) and subject (30).
```{r, max.print = 15}
meansdData <- data.frame(completeData[,1:2], X_meansd) %>%
              tbl_df() %>%
              group_by(Activity, Subject)%>%
              summarize_all(mean) %>%
              print
```

Details about the session
```{r}
sessionInfo()
```